---
title: "Task21.rmd"
author: "Brooks Groharing"
date: "11/6/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
library(tidyquant) # to get stock data
library(tidyverse) # for pipes
library(dplyr) # for data transformations
library(lubridate) # for date/time functions
library(timetk) # for converting dates to xts
library(dygraphs) # for interactive plots

#KR: stock for Kroger
#Load it into a tibble
df_kr = tq_get("KR",get="stock.prices")

xts = df_kr %>%
  pivot_wider(names_from=symbol,values_from=adjusted) %>% 
  filter(year(date)>2015) %>%
  select(date,KR) %>% tk_xts(data_var=date)

xts_monthly = df_kr %>% filter(date >= "2018-4-5") %>%
  tq_transmute(mutate_fun=monthlyReturn,col_rename=c("Monthly Return")) %>%
  tk_xts()


start = "2020-3-29"
present = "2020-11-04"

```

### Kroger Stock Performance, 5 yr

```{r echo=FALSE}

#Stock price performance over 5 years.
dygraph(data=xts,main="Kroger stock price performance, 2016-present")

```
### April 2018-Present

```{r echo=FALSE}

dygraph(data=xts_monthly,main="Kroger $10k stock performance, by month 
(4/5/18-present)") %>%
  dyRebase(value=10000) %>%
  dySeries(color="blue") %>%
  dyAnnotation("2019-3-29","A",tooltip="March 2019: KR stock falls 16 percent following dissappointing Q4 sales") %>%
  dyAnnotation("2020-3-31","B",tooltip="March 2020: The stock market crashes, signalling the start of an ongoing recession") %>%
  dyRangeSelector(dateWindow=c("2018-3-29","2020-11-04"),keepMouseZoom=TRUE)

```

This graph shows how a $10,000 investment in KR stock in April 2018 would have performed in the time since. I've highlighted two important times/trends in the data. First, KR stock prices underwent a broad downward trend throughout 2019, with especially poor quarterly sales discouraging investors in April (A). Stock continued to rise and fall through the remainder of the year. 2020 disrupted this pattern, with the pandemic and associated recession depressing stock prices from March onwards(B).

