---
title: "Task6: World Data Investigation"
author: "Brooks Groharing"
date: "9/15/2020"
output: 
  html_document: 
    keep_md: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE,echo=FALSE,code = readLines("task6.r")}
library(scales)
library(dplyr)
library(directlabels)
library(tidyverse)
library(ggrepel)


#devtools::install_github("drsimonj/ourworldindata")

decline = read_csv("economic-decline-in-the-second-quarter-of-2020.csv")
names(decline)[3] = "Growth"

filterlist=c('China','NAFTA','OECD - Total','EU','G7')

for (place in filterlist){
  print(place)
  decline = filter(decline,Entity != place)
}

decline = decline %>% arrange(Growth) %>%transform(Entity = paste(Entity," ",round(Growth,digits=1),"% ",sep=""),gradient=Growth)

titletext = "Economic decline in the second quarter of 2020"
subtext ="The percentage decline of GDP relative to the same quarter in 2019. It is adjusted for inflation."
captiontext = "Source: Eurostat,OECD and individual national statistics agencies."

bottomColor = rgb(247,239,71,maxColorValue = 255)
topColor = rgb(49,45,123,maxColorValue = 255)

#Graphing
Economy = ggplot(data=decline,aes(x=reorder(Entity,Growth),y=Growth)) +
  geom_col(aes(fill=gradient)) +
  scale_y_continuous() +
  coord_flip() +
  labs(title=titletext,subtitle=subtext,caption=captiontext) +
  geom_text(aes(y=Growth,hjust="right",label=reorder(Entity,Growth))) +
  theme_minimal() + 
  theme(
    panel.grid.minor=element_line(linetype = "dashed"),
    panel.grid.major=element_line(linetype = "dashed"),
    panel.grid.minor.y=element_blank(),
    panel.grid.major.y=element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    legend.position = "none")+
  scale_fill_gradient(low=bottomColor,high=topColor)


df2 = ourworldindata::child_mortality
mort2015 = df2 %>% filter(year == 2015) %>% arrange(child_mort) %>% filter(is.na(continent)==FALSE) %>% filter(is.na(child_mort)==FALSE)

top20names = head(mort2015,20)
bottom20names = tail(mort2015,10)

top20 = df2 %>% filter(country %in% top20names$country)
bottom20 = df2 %>% filter(country %in% bottom20names$country)

mortality = ggplot(data=top20,mapping=aes(x=year,y=child_mort)) + 
  geom_smooth(aes(group=country),color="orange") + geom_smooth(data=bottom20,aes(group=country),color='blue') +
  xlab("Year") + ylab("Child Mortality (log scale)") + ggtitle("Historical Child mortality in 10 highest and lowest states, as of 2015") +
  xlim(1800,2020) +
  ylim(0,750) + scale_y_log10() + theme_bw()
```


---
```{r}
Economy
```
This graphic shows how much each country's economy has suffered in the past year. The scale on the y-axis represents percentage growth relative to each country's economy, since 2019. With the notable exception of China (not pictured), all nations with recorded data in the dataset experienced negative groowth.


source: https://ourworldindata.org/covid-health-economy
---

```{r}

mortality

```
This second graph captures the trends in deaths under the age of five around the globe, using a historical dataset that goes back to around 1800 (earlier in a few countries). Since there were so many countries and not much space, I elected to depict only the top (orange) and bottom (blue) 20 countries in my graphic, to maintain readability while also showing the disparty between the poles. I also chose to use a logarithmicscale, for easier differentiation of the lines, and I used a a smoothing function to cut down on the (considerable) noise in the dataset. 