---
title: "Task6: World Data Investigation"
author: "Brooks Groharing"
date: "9/15/2020"
output: 
  html_document: 
    keep_md: yes
    fig_width: 10
    fig_height: 7
---
```{r setup, include=FALSE}
print("beep")
```


## Our World in Data Recreation
```{r include=FALSE,echo=FALSE}

library(scales)
library(dplyr)
library(directlabels)
library(tidyverse)

#Get dataset
decline = read_csv("economic-decline-in-the-second-quarter-of-2020.csv")
names(decline)[3] = "Growth"

filterlist=c('China','NAFTA','OECD - Total','EU','G7')

for (place in filterlist){
  print(place)
  decline = filter(decline,Entity != place)
}

decline = decline %>% arrange(Growth) %>%transform(Entity = paste(Entity," ",round(Growth,digits=1),"% ",sep=""),gradient=Growth)

#Formatting variables
titletext = "Economic decline in the second quarter of 2020"
subtext ="The percentage decline of GDP relative to the same quarter in 2019. It is adjusted for inflation."
captiontext = "Source: Eurostat,OECD and individual national statistics agencies."

bottomColor = rgb(247,239,71,maxColorValue = 255)
topColor = rgb(49,45,123,maxColorValue = 255)

#Building graph
Economy = ggplot(data=decline,aes(x=reorder(Entity,Growth),y=Growth)) +
  geom_col(aes(fill=gradient)) +
  scale_y_continuous() +
  coord_flip() +
  labs(title=titletext,subtitle=subtext,caption=captiontext) +
  geom_text(aes(y=Growth,hjust="right",label=reorder(Entity,Growth))) +
  theme_minimal() + 
  theme(
    panel.grid.minor=element_line(linetype = "dashed"),
    panel.grid.major=element_line(linetype = "dashed"),
    panel.grid.minor.y=element_blank(),
    panel.grid.major.y=element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    legend.position = "none")+
  scale_fill_gradient(low=bottomColor,high=topColor)
```

```{r echo=FALSE}
Economy
```




This graphic is reproduced from one featured in an Our World in Data report on the effect that COVID-19 and the recession have had on different countries' economies, and how this related to the scale of their health response. The scale on the y-axis represents percentage growth of each country's economy since 2019. With the exception of China (not pictured), all nations recorded in the original dataset experienced negative growth.

<font size="2">[Link to Source Data](ourworldindata.org/covid-health-economy)</font>


## Childhood Mortality Visualization
```{r include=FALSE,echo=FALSE}
#get datasets
df2 = ourworldindata::child_mortality
mort2015 = df2 %>% filter(year == 2015) %>% arrange(child_mort) %>% filter(is.na(continent)==FALSE) %>% filter(is.na(child_mort)==FALSE)

top20names = head(mort2015,20)
bottom20names = tail(mort2015,10)

top20 = df2 %>% filter(country %in% top20names$country)
bottom20 = df2 %>% filter(country %in% bottom20names$country)

#Build plot
mortality = ggplot(data=top20,mapping=aes(x=year,y=child_mort)) + 
  geom_smooth(aes(group=country),color="orange") + geom_smooth(data=bottom20,aes(group=country),color='blue') +
  xlab("Year") + ylab("Child Mortality (log scale)") + ggtitle("Historical Child mortality in 10 highest and lowest states, as of 2015") +
  xlim(1800,2020) +
  ylim(0,750) + scale_y_log10() + theme_bw()


```

```{r echo=FALSE,results='hide',message=FALSE,warning=FALSE}
#Why is rmd so determined to print console stuff to file...
mortality
```


This second graph captures the trends in deaths under the age of five by country, using a historical dataset that goes back to around 1800 (and earlier, in a few placess). Since there are way too many countries in the dataset to include in one plot, I elected to depict only the top (orange) and bottom (blue) 20 countries in my graphic. The idea was to maintain readability, while also emphasizing the disparity in mortality rates throughout the world. I used a smoothing function and a logarithmic scale, again to cut down on clutter and make the overall trend in each country more readable.