---
title: 'Task 5: Data Import and ggplot2'
author: "Brooks Groharing"
date: "9/10/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

### What I learned
I learned:
  -How to download online datasets and libraries in code
  -How to import .csv files, and how R tries to interpret their values 
  -More options in ggplot, like blanking out elements and adding offset labels with ggrepel
  -More options for graph themes and color schemes
  
### What I found difficult
I had some difficulty importing the dataset from 538 (below).

```{r include = FALSE, collapse=TRUE}
library(tidyverse)
```

## FiveThirtyEight Data

The dataset I grabbed is from 538's "Tracking Congress In The Age Of Trump," a live page tracking how often each member of congress votes in line with the President, and ranks them all.

Link to webpage: https://projects.fivethirtyeight.com/congress-trump-score/
Link to datasets: https://github.com/fivethirtyeight/data/tree/master/congress-trump-score

I downloaded just the averagess.csv dataset for this assignment, and ignored the voting predictions data.

### Importing it

```{r results='hide'}
download.file("https://projects.fivethirtyeight.com/congress-tracker-data/csv/averages.csv","congress-tracking.csv")

df = read_csv("congress-tracking.csv")
```

### Data Manipulations

First I consolidated individual politicians:
```{r,collapse=TRUE}
politicians = summarise(group_by(df,last_name,chamber,state,district), avg = mean(agree_pct,na.rm=TRUE))
```
Then, I averaged their "Trump scores" by state.
```{r,collapse=TRUE}
states = summarise(group_by(politicians,state),mean(avg,na.rm=TRUE))
```

### Saving Transformed Data
Lastly, I saved the aggregated state data as a .csv.

```{r}
write.csv(states,"AgreementScoreByState.csv")
```