---
title: "RForDataScienceNotes"
author: "Brooks Groharing"
date: "9/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
URL: 
https://r4ds.had.co.nz/transform.html


```{r include = FALSE}
library(nycflights13)
library(tidyverse) #overwrites states::filter()
```
nycflights13 includes NYC flights, from 2013

tibble- special type of df tweaked for tidyverse

| type | desc |
| --- | ----- |
| int |integer|
| dbl |double |
| chr |string |
| dttm |date/time|
| lgl |boolean|
| fctr |factor(categorical var w/ fixed levels)|
| date |...date|


## Operators
OR: |
AND: &
NOT: !()

### filtering

gets rows that fit given criteria

filter() only returns NA if requested (is.na())


(filter(flights,month==1,day==1))

filter(flights, !month == 11 && day == 1)
nov_dec <- filter(flights, month %in% c(11, 12))


---

### arrange()

arrange() orders rows by given column(s)

syntax: arrange(flights,month,desc(day))

---

### select()
select gets specific columns

syntax: select(flights,year,month,Day = day)

### mutations

Mutate to compute new columns and append;
transmute() only uses new variables

```{r}
flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)

transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```

also useful- mean(), sum()

%% remainder

%/% integer division

^ for powers

log(),log2(),log10()

---
### summarise()

summarise() gives summary data
pair with group_by()

REMEMBER- na.rm is needed in aggregates to ignore missing values; most operations on na will return na.

```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))

by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

### Pipe %>%
Use the pipe to run a series of commands, without intermediate steps cluttering everything. ex:

```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
```

### Stats stuff

various useful functions: 

min(x)

max(x)

sd(x)

n(x)

n_distinct(x)

sum(!is.na(x))

count(x)